Sale difference between Loblaws and Voila

### 2024-11-15 paper/paper.qmd

## Overview
We use the statistical programming language R [@citeR].... The data used in this paper came from the [@hammer]. Following @tellingstories, we considerto capture the pricing dynamics of Ambrosia apples across different vendors, we extracted relevant data from a larger dataset containing multiple products and vendors. The initial dataset included information from seven vendors, each offering various products. Our focus was on the specific product, "Ambrosia Apples," which was sold by two vendors—NoFrills and Loblaws.From this comprehensive dataset, we filtered out all other products, isolating only the entries related to Ambrosia apples. The extracted data included key information such as the product ID, product name, current price, and old price. Additionally, a field labeled "other" contained further notes on sale conditions or discounts. The dataset also recorded timestamps, allowing us to track pricing changes over time.

@@

	1.	Price Comparison Between Vendors:
	•	NoFrills offers a current price for the apples ranging from $5.49 to $5.99, with an old price of $6.49 for all sale items. This indicates a $1.00 decrease in price during the sale period, suggesting that NoFrills is promoting a sale by reducing the price from a regular price point.
	•	Loblaws, on the other hand, has a consistent current price of $5.49 for all data points, which is generally lower than the sale price at NoFrills. Additionally, Loblaws’ old price fluctuates slightly around $5.49 to $6.58, but the regular price appears stable without the promotional fluctuations observed at NoFrills.
	2.	Sale Behavior Analysis at NoFrills:
	•	At NoFrills, the old price for the apples is consistently set at $6.49 before being reduced to $5.49 during sale periods. This consistent pricing behavior suggests that NoFrills applies a fixed discount strategy, where the regular price is consistently reduced by $1.00 during promotional events.
	•	The stable difference between the sale and non-sale prices raises questions about whether NoFrills may be inflating prices just before the sale period to create the perception of a significant discount. The predictable pattern of a $1.00 reduction during sales could indicate a structured approach to pricing, potentially aimed at attracting price-sensitive customers through a visible discount.
	3.	Price Consistency at Loblaws:
	•	Loblaws displays a more stable pricing pattern, with the current price consistently at $5.49 and only slight variation in old prices. Unlike NoFrills, Loblaws does not engage in visible promotional adjustments, indicating a transparent and steady pricing strategy that does not rely on artificial price fluctuations.
	•	The absence of sale-driven price changes suggests that Loblaws might prioritize a straightforward pricing model that could appeal to customers who value consistency over promotional sales.
	4.	Conclusion on Price Manipulation:
	•	The analysis indicates that NoFrills may engage in a pricing strategy where prices are elevated to $6.49 before sales and then reduced back to a sale price of $5.49, giving the impression of a more substantial discount. This strategy could be perceived as artificial price manipulation, potentially intended to create a psychological effect on customers.
	•	In contrast, Loblaws maintains a consistent price, which may be perceived as more transparent and consumer-friendly, as it avoids the potential illusion of savings through price fluctuations.

@@

# Conclusion

In conclusion, the analysis suggests that NoFrills may utilize price elevation tactics before sales to create an illusion of larger discounts, while Loblaws maintains a stable and transparent pricing model. Understanding these pricing strategies allows consumers to make more informed purchasing decisions. Further research into broader pricing practices would enhance the insights from this analysis, potentially examining consumer responses to these strategies.

>>>>

## Overview
We use the statistical programming language R [@citeR]. The data used in this paper came from the [@hammer]. Following @tellingstories, we considerto capture the pricing dynamics of Ambrosia apples across different vendors, we extracted relevant data from a larger dataset containing multiple products and vendors. The initial dataset included information from seven vendors, each offering various products. Our focus was on the specific product, "Ambrosia Apples," which was sold by two vendors—NoFrills and Loblaws.From this comprehensive dataset, we filtered out all other products, isolating only the entries related to Ambrosia apples. The extracted data included key information such as the product ID, product name, current price, and old price. Additionally, a field labeled "other" contained further notes on sale conditions or discounts. The dataset also recorded timestamps, allowing us to track pricing changes over time.

@@

## Price Comparison Between Vendors
Loblaws offers a current price for the apples \$1.21, with an old price of \$1.45 for all sale items. This suggests a $0.24 decrease in price during the sale period, indicating that Loblaws is promoting a sale by reducing the price from the regular price.
NoFrills, on the other hand, has a consistent current price of $1.2 for all data points. This is the same as the sale price at Loblaws. Additionally, NoFrills has no old price since they didn't do any discounts. However, the regular price at NoFrills appears to be stable without the promotional fluctuations observed at Loblaws.

## Conclusion: Price Manipulation Practices?
Despite Loblaws advertising a price drop from \$1.45 to \$1.21 as part of a "sale," the price at NoFrills is consistently set at \$1.21, without any promotional claims. This raises suspicions that Loblaws may be engaging in price manipulation practices, where the advertised sale price is simply a temporary reduction to match NoFrills' regular price, rather than offering a genuine discount.
These findings highlight potential concerns about transparency in pricing and suggest that consumers may be misled into believing they are receiving a larger discount than is actually being offered. Loblaws' "sale" price may not reflect a true reduction in value, but rather a strategic pricing move to attract customers by promoting the illusion of a better deal. This could be seen as a form of pricing manipulation, where the apparent discount is used to create the perception of a more significant bargain.
Consumers should be cautious when interpreting sales promotions and be aware of the actual price comparisons across different vendors to make informed purchasing decisions.

@@

## Correlation vs. Causation
In this analysis, we have observed a correlation between the sale price at Loblaws and the regular price at NoFrills, suggesting that both prices are set at $1.21 during the sale period. However, it is important to note that correlation does not imply causation. While the two prices may appear to be related, this analysis does not establish a causal relationship between Loblaws' price reduction and NoFrills' pricing strategy. For example, it is possible that Loblaws adjusted its price temporarily to match NoFrills' pricing rather than making a true price reduction. A more comprehensive causal analysis would require additional data points, such as customer behavior during sale periods, or experiments where prices are manipulated independently, to determine whether Loblaws' pricing strategy is indeed a response to NoFrills or if there are other factors influencing both vendors' pricing.

## Missing Data
Our dataset consists of prices from two vendors (NoFrills and Loblaws), but the full dataset includes many more products and vendors. Missing data could be a potential concern in our analysis, especially if key information, such as prices for other vendors, was not available or if data points for certain time periods were omitted. Missing data can bias the analysis if it is not random; for example, if sales data is missing during peak shopping periods when prices are likely to be highest, this could result in a lower estimate of the actual price difference. In our case, the absence of historical prices for NoFrills or detailed pricing data for other vendors might limit the conclusions we can draw about the broader pricing landscape. To address this, we could perform sensitivity analyses to evaluate how the results change with different assumptions about the missing data, or use imputation techniques to fill in gaps where appropriate.

## Sources of Bias
There are several potential sources of bias in this study that could influence our results. First, selection bias could occur if the data points from Loblaws and NoFrills are not representative of all pricing periods, or if certain time frames (e.g., weekends or holidays) are overrepresented. Since our data focuses on a subset of vendors and products, it may not fully capture the diversity of pricing practices across all 7 vendors in the full dataset. Additionally, vendor-specific biases may also exist if either Loblaws or NoFrills have particular promotional strategies that differ from those of other vendors. For instance, Loblaws' advertised sales may be selectively applied to high-demand products like Ambrosia apples, while NoFrills might use a more consistent pricing approach across its entire range of produce. Furthermore, potential reporting biases could affect the accuracy of the data if certain pricing changes were not accurately recorded or if vendor promotions were not consistently reflected in the dataset. To mitigate these biases, future analyses could involve a broader set of vendors and products, as well as methods to ensure that the dataset is representative and balanced.
